Non-parametric approaches to optimal policy are more robust

### 2013-03-29 inst/doc/nonparametric-bayes.Rmd
automate bibliographic extraction

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological interactions
involved, in practice real mechanisms are often unknown and the true
dynamics too complex to be captured by simple models [@refs]. While
simple mechanistic models can nevertheless provide imporant insights
into possible dynamics -- for instance, demonstrating that a vaccine
does not have to be 100% effective to eliminate the transmission of a
virus [@Kermack1921] -- such approaches are not well-suited for use in
forecasting outcomes of potential management options.  Non-parametric 
approaches offer a more flexible alternative that can both more accurately
reflect the data available while also representing greater uncertainty
in areas (of state-space) where data is lacking.  Ecological research
and management strategy should pay closer attention to the opportunities 
and challenges nonparametric modeling can offer.

@@

We argue that while simple mechanistically motivated models
may be best for the latter case [@Cunningham2013; @Geritz2012], such models 
can be not only inaccurate but misleading in quantitative decision making.
Nonparametric models can more realistically represent uncertainties outside 
of observed data while also better capturing the dynamics in the region
observed.  Advances in the theory and computational implementations of 
nonparametric methods make them ripe for such applications.  We use the classic
problem of optimal harvest of a marine fishery to illustrate how the 
nonparametric approach of Gaussian processes can be applied.

@@

### The fisheries management problem

@@

Given parameters for the function $f$ and probability distribution $Z$, 
along with a given economic model determining the  price/profit $\Pi(X_t, h_t)$
realized in a given year given a choice of harvest $h_t$ and observed stock $X_t$.
This problem can be solved exactly for discretized values of stock $X$ and policy $h$
using stochastic dynamic programming [@Mangel1982].  
Problems of this sort underpin much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of $f$,
along with its parameters.  The standard approach [@refs] uses one of a 
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any 
available existing data. Though modifications of this approach mentioned in the 
introduction can permit additional sources of uncertainty such as measurement 
error in the stock assessment, implementation errors in the harvest policy, [@Clark1986, @Roughgarden1996, @Sethi2005]
uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012]

Here, we compare this approach to our alternative
that uses a Gaussian Process in place of the a given stock recruitment curve.

@@

The use of Gaussian process regression (known as Kreging in the geospatial
literature) to formulate a predictive model is relatively new in the
context of modeling dynamical systems [@Kocijan2005] and introduced
in the ecological modeling and fisheries management by @Munch2005.
An accessible and thorough introduction to the formulation and use of
Gaussian processes can be found in @Rasmussen2006.

The essense of the Gaussian process approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
Gaussian process approach can be thought of as a computationally
efficient approximation to this approach. Gaussian processes represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the Gaussian
Process approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<!-- Figure 1 include curves drawn from the posterior density?  -->

Following [@Munch2005], we will estimate Gaussian processes from

@@

###

>>>>

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological interactions
involved, in practice real mechanisms are often unknown and the true
dynamics too complex to be captured by simple models [@refs]. While
simple mechanistic models can nevertheless provide imporant insights
into possible dynamics -- for instance, demonstrating that a vaccine
does not have to be 100% effective to eliminate the transmission of a
virus  -- such approaches are not well-suited for use in
forecasting outcomes of potential management options.  Non-parametric 
approaches offer a more flexible alternative that can both more accurately
reflect the data available while also representing greater uncertainty
in areas (of state-space) where data is lacking.  Ecological research
and management strategy should pay closer attention to the opportunities 
and challenges nonparametric modeling can offer.

@@

We argue that while simple mechanistically motivated models
may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
can be not only inaccurate but misleading in quantitative decision making.
Nonparametric models can more realistically represent uncertainties outside 
of observed data while also better capturing the dynamics in the region
observed.  Advances in the theory and computational implementations of 
nonparametric methods make them ripe for such applications.  We use the classic
problem of optimal harvest of a marine fishery to illustrate how the 
nonparametric approach of Gaussian processes can be applied.

@@

### The optimal control problem in fisheries management

@@

Given parameters for the function $f$ and probability distribution $Z$, 
along with a given economic model determining the  price/profit $\Pi(X_t, h_t)$
realized in a given year given a choice of harvest $h_t$ and observed stock $X_t$.
This problem can be solved exactly for discretized values of stock $X$ and policy $h$
using stochastic dynamic programming (SDP) [@Mangel1982].  
Problems of this sort underpin much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of $f$,
along with its parameters.  The standard approach [@refs] uses one of a 
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any 
available existing data. Though modifications of this approach mentioned in the 
introduction can permit additional sources of uncertainty such as measurement 
error in the stock assessment, implementation errors in the harvest policy, [@Clark1986, @Roughgarden1996, @Sethi2005]
uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012],
we focus on the simplest example to illustrate the comparison, where only a 
single form of uncertainty (the growth shocks $Z_t$) is present in the underlying
model, making additional modes unnecessary. 
<!--Add comment / apology on single species dynamics -->

We compare this approach to our alternative that uses a Gaussian Process (GP)
in place of the a given stock recruitment curve.  We briefly revisit
a description of the Gaussian Process approach that has appeared in 
recent literature before describing how we take the important step of 
adapting the GP to the decision theory problem that can still be solved
using SDP.

@@

The use of Gaussian process (GP) regression (known as Kreging in the geospatial
literature) to formulate a predictive model is relatively new in the
context of modeling dynamical systems [@Kocijan2005] and introduced
in the ecological modeling and fisheries management by @Munch2005.
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

The essense of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<!-- Figure 1 include curves drawn from the posterior density?  -->

@@

### 2013-03-29 inst/doc/nonparametric-bayes.tex
automate bibliographic extraction

\begin{document}
\begin{frontmatter}
  \title{Non-parametric approaches to optimal policy are more robust}
  \author[cstar]{Carl Boettiger\corref{cor1}}
  \author[cstar]{Marc Mangel}
  \author[noaa]{Stephan B. Munch}
  \ead{cboettig@ucsc.edu}
  \cortext[cor1]{Corresponding author, cboettig@ucsc.edu}
  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 \end{frontmatter}

@@

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological
interactions involved, in practice real mechanisms are often unknown and
the true dynamics too complex to be captured by simple models
{[}@refs{]}. While simple mechanistic models can nevertheless provide
imporant insights into possible dynamics -- for instance, demonstrating
that a vaccine does not have to be 100\% effective to eliminate the
transmission of a virus {[}@Kermack1921{]} -- such approaches are not
well-suited for use in forecasting future dynamics upon which outcomes
management policy can be based. Despite a long history of work
emphasizing the difference between modeling for understanding (generic
or strategic models) and modeling for prediction (precise, specific, or
tactical models) (e.g. Levins 1966), simple parametric models continue
to find application in predictive management contexts. The management of
marine fisheries is a prime example, in which simple, mechanistically
motivated models of stock-recruitment relationships such as Ricker or
Beverton-Holt curves are systematically fit to data and used in solving
decision-theoretic or optimal control problems determining resource
management policies such as maximum sustainable yield {[}@refs{]}.

Stochastic dynamic programming @Mangel1982

To address structural uncertainty in the processes involved, such
mechanistic models can be adapted to accomidate uncertainty in the
parameters or consider a set of alternative models simultaneously
(belief SDP)

Further unknowns such as measurement uncertainty, parameter uncertainty,
unobserved states, knowledge of boundary conditions, etc. further
compound the issue. Though a hierarchical Bayesian approach provides a
natural way to address these from a statistical standpoint, formulating
reasonable parametric descriptions of each form of uncertainty is a
challenging task in itself, let alone the computational difficulties of
solving such a system. Cressie et al. (2009) provides a good account of
the successess and challenges of the approach. Applying these approaches
in the management context of sequential decision making, in which
forecasts must be obtained over a range of possible actions and updated
regularly as new information arrives makes such an approach less
feasible still.

Though machine learning approaches have begun to appear in the
ecological and conservation literature (Species distribution models),
including the Gaussian process based approach used here (Munch et al.
2005), they remain unfamiliar and untrusted approaches for most
ecologists. Machine learning approaches represent an essentially
pattern-based rather than process-based approach, raising the same
skepticism from most theoretical ecologists that they hold for older
correlative methods such as linear regression, while the complexity of
the techniques has barred their adoption in more empirical audiences.

A more immediate barrier to their adoption is the absence of a framework
for applying machine learning approaches to resource management
problems. Traditional approaches to optimal control (Pontryagin's
principle, stochastic dynamic programming) rely on knowledge of the
state equation, usually described by a simple parametric model. Here we
illustrate how a stochastic dynamic programming algorithm can
alternatively be driven by the predictions from a Gaussian process -- a
machine learning approximation to the state dynamics.\\

@@

@@

The use of Gaussian process regression (known as Kreging in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems {[}@Kocijan2005{]} and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of Gaussian processes can be found in
@Rasmussen2006.

The essense of the Gaussian process approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters; or
using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
Gaussian process approach can be thought of as a computationally
efficient approximation to this approach. Gaussian processes represent a
large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection. By modeling at the
level of the process, rather than the level of parametric equation, we
can more concisely capture the possible behavior of these curves. In
place of a parametric model of the dynamical system, the Gaussian
Process approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us to
consider a set of possible curves simultaneously.\\

@@

Predictive accuracy is not the goal of all modeling, as ecologists have
been observing for as long as they made models (perhaps none more
memorably than @Levins1969). Mechanistic modeling is at its most
powerful not when it is used to directly forecast future states but when
it provides an understanding of how to approach a problem. SIR-type
models from the epidemiological literature are a good example. While the
simplest SIR models have little predictive power over the outbreak
intensity or timing at a particular location, they provide a powerful
understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate
every member of the population to stop the spread, but rather it
suffices to vaccinate a sufficient fraction of the population to reduce
$R_0$ below 1.

@@

The posterior Gaussian process is completely defined by the expected
value and covariance matrix at a defined set of training points. For
simplicty we will consider a these points to fall on a discrete, uniform
grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
$K$). Again to keep things simple we will use this same grid
discritization for the parametric approach. Other options for choosing
the grid points, including collocation methods and functional basis
expansion (or even using Guassian processes in place of the discrete
optimization; an entirely different context in which GP can be used in
SDP, see {[}@Deisenroth2009{]}) could also be considered.

@@

@Gramacy2005

@@

Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
2005. ``Population extinction in discrete-time stochastic population
models with an Allee effect.'' \emph{Journal of Difference Equations and
Applications} 11 (4-5) (apr): 273--293.
doi:10.1080/10236190412331335373.

Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.

Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
Effects in Ecology and Conservation}. Oxford University Press, USA.

Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
Christopher K. Wikle. 2009. ``Accounting for uncertainty in ecological
analysis: the strengths and limitations of hierarchical statistical
modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.

Levins, Richard. 1966. ``The strategy of model building in population
biology.'' \emph{American Scientist} 54 (4): 421--431.

Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
Behavioral Ecology}. Princeton: Princeton University Press.

Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
691--695. doi:10.1111/j.1461-0248.2005.00766.x.

Reed, William J. 1979. ``Optimal escapement levels in stochastic and
deterministic harvesting models.'' \emph{Journal of Environmental
Economics and Management} 6 (4) (dec): 350--363.
doi:10.1016/0095-0696(79)90014-7.

Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
(6856) (oct): 591--6. doi:10.1038/35098000.

>>>>

\begin{document}
\begin{frontmatter}
  \title{Non-parametric approaches to optimal policy are more robust}
  \author[cstar]{Carl Boettiger\corref{cor1}}
  \cortext[cor1]{Corresponding author}
  \ead{cboettig@ucsc.edu}
  \author[cstar]{Marc Mangel}
  \author[noaa]{Stephan B. Munch}
  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 \end{frontmatter}

@@

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological
interactions involved, in practice real mechanisms are often unknown and
the true dynamics too complex to be captured by simple models
{[}@refs{]}. While simple mechanistic models can nevertheless provide
imporant insights into possible dynamics -- for instance, demonstrating
that a vaccine does not have to be 100\% effective to eliminate the
transmission of a virus -- such approaches are not well-suited for use
in forecasting outcomes of potential management options. Non-parametric
approaches offer a more flexible alternative that can both more
accurately reflect the data available while also representing greater
uncertainty in areas (of state-space) where data is lacking. Ecological
research and management strategy should pay closer attention to the
opportunities and challenges nonparametric modeling can offer.

\subsubsection{Quantitative vs Qualitative Decisions}

In this paper, we consider those ecological management problems in which
a mathematical (or computational) model is used to quantitatively inform
decision-making by allowing a manager to compare to the expected
consequences of potential management actions (or policies). We
distinguish this from the solely qualitative use of a model, in which
models are used to represent and compare hypotheses of different
mechanisms that would lead to qualitatively different actions.

We argue that while simple mechanistically motivated models may be best
for the latter case (Cuddington, Fortin, and Gerber 2013; Geritz and
Kisdi 2012), such models can be not only inaccurate but misleading in
quantitative decision making. Nonparametric models can more
realistically represent uncertainties outside of observed data while
also better capturing the dynamics in the region observed. Advances in
the theory and computational implementations of nonparametric methods
make them ripe for such applications. We use the classic problem of
optimal harvest of a marine fishery to illustrate how the nonparametric
approach of Gaussian processes can be applied.

We imagine the manager seeks to maximize some known socio-economic value
function, given their knowledge of the system. In particular, we will
focus on the case in which the manager faces a series of decisions, such
as setting a harvest quota for a marine fishery each year based on
observations of the stock size the year before.

\subsubsection{Uncertainties: the classical approach}

The ecological dynamics of most systems of management interest are
typically both complex and unknown {[}@refs{]}. Despite this,
quantitative policy optimization in these systems has almost exclusively
been based on simple, mechanistically motivated parametric models. There
are several reasons for this. First, limited available data almost
always percludes any accurate estimate of more complicated models.
Second, the computational demands of determining an optimal stategy in a
sequential decision-theory problem suffer from the ``curse of
dimensionality'' that makes\\it difficult to consider more complicated
models, states, or actions. Unfortunately, these challenges also make it
particularly difficult to reflect the true uncertainty either in the
parameters of the given models (parametric uncertainty) or in the
structure of the models themselves (structural uncertainty).\\Further
unknowns such as measurement uncertainty, parameter uncertainty,
unobserved states, knowledge of boundary conditions, etc. further
compound the issue. Though a hierarchical Bayesian approach provides a
natural way to address these from a statistical standpoint, formulating
reasonable parametric descriptions of each form of uncertainty is a
challenging task in itself, let alone the computational difficulties of
solving such a system. Cressie et al. (2009) provides a good account of
the successess and challenges of the approach. Applying these approaches
in the management context of sequential decision making, in which
forecasts must be obtained over a range of possible actions and updated
regularly as new information arrives makes such an approach less
feasible still.

An active literature and growing computational power over the past
several decades have only marginally improved this situation. Parametric
or structural uncertainty can be introduced only by increasing the state
space to reflect a distribution of possible parameters {[}@refs{]} or
the degree of belief in each of a set of possible models {[}@refs{]}
Frequently, the space of possible actions must then be reduced or the
algorithms adjusted by approximations to keep computations feasible.

Though machine learning approaches have begun to appear in the
ecological and conservation literature (Species distribution models),
including the Gaussian process based approach used here (Munch et al.
2005), they remain unfamiliar and untrusted approaches for most
ecologists. One potential barrier to their adoption is the absence of a
framework for applying machine learning approaches to resource
management problems. Traditional approaches to optimal control
(Pontryagin's principle, stochastic dynamic programming) rely on
knowledge of the state equation, usually described by a simple
parametric model. Here we illustrate how a stochastic dynamic
programming algorithm can alternatively be driven by the predictions
from a Gaussian process -- a machine learning approximation to the state
dynamics.

@@

\subsubsection{The optimal control problem in fisheries management}

In our example, we focus on the problem in which a manager must set the
harvest level for a marine fishery each year to maximize the net present
value of the resource, given an estimated stock size from the year
before. Rich data and global concerns have made marine fisheries the
crucible for much of the founding work (Gordon 1954; Reed 1979; May et
al. 1979; Ludwig and Walters 1982) in managing ecosystems under
uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
2007; Worm et al. 2009) have made understanding these challenges all the
more pressing.

To permit comparisons against a theoretical optimum we will consider
data on the stock dynamics simulated from a simple parametric model in
which recruitment of the fish stock $X_{t+1}$ in the following year is a
stochastic process governed by a function $f$ of the current stock
$X_t$, selected harvest policy $h_t$, and noise process $Z$,

\[X_{t+1} = Z_t f(X_t, h_t) \]

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) {[}@Mangel1982{]}.\\Problems of this sort underpin
much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach {[}@refs{]} uses
one of a handful of common parametric models representing the
stock-recruitment relationship, usually after estimating the model
parameters from any available existing data. Though modifications of
this approach mentioned in the introduction can permit additional
sources of uncertainty such as measurement error in the stock
assessment, implementation errors in the harvest policy, (Clark and
Kirkwood 1986, , ) uncertainty in parameters (Mangel and Clark 1985,
@Schaupaugh2013) or model structure (B. K. Williams 2001, ), we focus on
the simplest example to illustrate the comparison, where only a single
form of uncertainty (the growth shocks $Z_t$) is present in the
underlying model, making additional modes unnecessary.

We compare this approach to our alternative that uses a Gaussian Process
(GP) in place of the a given stock recruitment curve. We briefly revisit
a description of the Gaussian Process approach that has appeared in
recent literature before describing how we take the important step of
adapting the GP to the decision theory problem that can still be solved
using SDP.

\subsubsection{The Non-parametric Bayesian alternative for
stock-recruitment curves}

The use of Gaussian process (GP) regression (known as Kreging in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in C. E. Rasmussen and Williams
(2006).

The essense of the GP approach can be captured in the following thought
experiment: An exhaustive parametric approach to the challenge of
structural uncertainty might proceed by writing down all possible
functional forms for the underlying dynamical system with all possible
parameter values for each form, and then consider searching over this
huge space to select the most likely model and parameters; or using a
Bayesian approach, assign priors to each of these possible models and
infer the posterior distribution of possible models. The GP approach can
be thought of as a computationally efficient approximation to this
approach. GPs represent a large class of models that can be though of as
capturing or reasonably approximating the set of models in this
collection. By modeling at the level of the process, rather than the
level of parametric equation, we can more concisely capture the possible
behavior of these curves. In place of a parametric model of the
dynamical system, the GP approach postulates a prior distribution of
(n-dimensional) curves that can be though of as approximations to a
range of possible (parametric) models that might describe the data. The
GP allows us to consider a set of possible curves simultaneously.\\

@@

The posterior Gaussian process is completely defined by the expected
value and covariance matrix at a defined set of training points. For
simplicty we will consider a these points to fall on a discrete, uniform
grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
$K$). Again to keep things simple we will use this same grid
discritization for the parametric approach. Other options for choosing
the grid points, including collocation methods and functional basis
expansion (or even using Guassian processes in place of the discrete
optimization; an entirely different context in which GP can be used in
SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
considered.

@@

Predictive accuracy is not the goal of all modeling, as ecologists have
been observing for as long as they made models (perhaps none more
memorably than Levins (1969)). Mechanistic modeling is at its most
powerful not when it is used to directly forecast future states but when
it provides an understanding of how to approach a problem. SIR-type
models from the epidemiological literature are a good example. While the
simplest SIR models have little predictive power over the outbreak
intensity or timing at a particular location, they provide a powerful
understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate
every member of the population to stop the spread, but rather it
suffices to vaccinate a sufficient fraction of the population to reduce
$R_0$ below 1.

@@

The posterior Gaussian process is completely defined by the expected
value and covariance matrix at a defined set of training points. For
simplicty we will consider a these points to fall on a discrete, uniform
grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
$K$). Again to keep things simple we will use this same grid
discritization for the parametric approach. Other options for choosing
the grid points, including collocation methods and functional basis
expansion (or even using Guassian processes in place of the discrete
optimization; an entirely different context in which GP can be used in
SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
considered.

@@

Gramacy (2005)

@@

Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
2005. ``Population extinction in discrete-time stochastic population
models with an Allee effect.'' \emph{Journal of Difference Equations and
Applications} 11 (4-5) (apr): 273--293.
doi:10.1080/10236190412331335373.
\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.

Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
control with uncertain stock dynamics: When, and how, to be
precautious.'' \emph{Journal of Environmental Economics and Management}
63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.

Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.

Clark, Colin W., and Geoffrey P. Kirkwood. 1986. ``On uncertain
renewable resource stocks: Optimal harvest policies and the value of
stock surveys.'' \emph{Journal of Environmental Economics and
Management} 13 (3) (sep): 235--244. doi:10.1016/0095-0696(86)90024-0.
\url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.

Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
Effects in Ecology and Conservation}. Oxford University Press, USA.
\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.

Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
Christopher K. Wikle. 2009. ``Accounting for uncertainty in ecological
analysis: the strengths and limitations of hierarchical statistical
modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.

Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
``Process-based models are required to manage ecological systems in a
changing world.'' \emph{Ecosphere} 4 (February): 1--12.
doi:10.1890/ES12-00178.1.
\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.

Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
``Gaussian process dynamic programming.'' \emph{Neurocomputing} 72 (7-9)
(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.

Geritz, Stefan a. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
mechanistic models?'' \emph{Journal of mathematical biology} 65 (6-7)
(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.

Gordon, H. S. 1954. ``The economic theory of a common-property resource:
the fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
\url{http://www.jstor.org/stable/10.2307/1825571}.

Gramacy, Robert B. 2005. ``BAYESIAN TREED GAUSSIAN PROCESS MODELS.''

Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and their
Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
doi:10.1007/s10021-007-9100-5.
\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.

Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
Murray-Smith. 2005. ``Dynamic systems identification with Gaussian
processes.'' \emph{Mathematical and Computer Modelling of Dynamical
Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.

Levins, Richard. 1969. ``Some demographic and genetic consequences of
environmental heterogeneity for biological control.'' \emph{Bulletin of
the Entomological Society of America} 15 (3): 237--240.
\url{http://www.ingentaconnect.com/content/esa/besa/1969/00000015/00000003/art00009}.

Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal harvesting with
imprecise parameter estimates.'' \emph{Ecological Modelling} 14 (3-4)
(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.

Mangel, Marc, and Colin W. Clark. 1985. \emph{Decision and control in
uncertain resource systems, Volume 172 (Mathematics in Science and
Engineering)}. Academic Press.
\url{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202}.

---------. 1988. \emph{Dynamic Modeling in Behavioral Ecology}.
Princeton: Princeton University Press.

May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
R. M. Laws. 1979. ``Management of multispecies fisheries.''
\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
doi:10.1126/science.205.4403.267.
\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.

Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
691--695. doi:10.1111/j.1461-0248.2005.00766.x.
\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.

Rasmussen, C. E., and C. K. I. Williams. 2006. \emph{Gaussian Processes
for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT Press,.
\url{www.GaussianProcess.org/gpml}.

Reed, William J. 1979. ``Optimal escapement levels in stochastic and
deterministic harvesting models.'' \emph{Journal of Environmental
Economics and Management} 6 (4) (dec): 350--363.
doi:10.1016/0095-0696(79)90014-7.
\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.

Roughgarden, Joan E., and F. Smith. 1996. ``Why fisheries collapse and
what to do about it.'' \emph{Proceedings of the National Academy of
Sciences of the United States of America} 93 (10): 5078.
\url{http://www.pnas.org/content/93/10/5078.short}.

Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
(6856) (oct): 591--6. doi:10.1038/35098000.
\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.

Sethi, Gautam, Christopher Costello, Anthony Fisher, Michael Hanemann,
and Larry Karp. 2005. ``Fishery management under multiple uncertainty.''
\emph{Journal of Environmental Economics and Management} 50 (2) (sep):
300--318. doi:10.1016/j.jeem.2004.11.005.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069605000057}.

Williams, Byron K. 2001. ``Uncertainty , learning , and the optimal
management of wildlife.'' \emph{Environmental and Ecological Statistics}
8: 269--288. doi:10.1023/A:1011395725123.

Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
of biodiversity loss on ocean ecosystem services.'' \emph{Science (New
York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.

Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
``Rebuilding global fisheries.'' \emph{Science (New York, N.Y.)} 325
(5940) (jul): 578--85. doi:10.1126/science.1173146.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
